SmartCheck

### 2012-11-10 paper/paper.tex
paper updates.

@@

@@

@@

>>>>

\newcommand{\ttp}[1]{\texttt{#1}}

@@

The QuickCheck testing framework was a revolutionary step-forward in
type-directed random testing~\cite{qc}.  Nowadays, QuickCheck is a standard
stage of a program development cycle in Haskell.

Because QuickCheck generates random values for testing, counterexamples it finds
may not be minimal.  In their original QuickCheck paper~\cite{qc}, the authors
report the following user experience by Andy Gill:
%
\begin{quote}
Sometimes the counter examples found are very large and it is difficult to go
back to the property and understand why it is a counter example.
\end{quote}
%
\noindent
QuickCheck defines a type class \ttp{Arbitrary} that defines a function
\ttp{arbitrary} for generating random values of a given type.  Gill added another
method to the type class:
%
\begin{code}
smaller :: a -> [a] 
\end{code}
%
\noindent
The purpose of \ttp{smaller} is to generate strictly smaller values (according to
some measure) from a given counterexample to attempt to find a smaller
counterexample.  Today, \ttp{smaller} is called \ttp{shrink}.

One of the dirty little secrets of QuickCheck is that even for simple data types
and simple properties, counterexamples can be much larger than the minimal
counterexamples.  Intuitively, there are two reasons: (1) because values are
generated randomly, a smaller value that produces a counterexample is skipped;
and (2) for a composite type, larger values are generated in tandem.  For
example, consider the following data type and property:
%
\begin{code}
data A = A Int deriving Show

instance Arbitrary A where
  arbitrary = liftM A arbitrary

prop :: [A] -> Property
prop x = property $ length x < 5
\end{code}
%
\noindent
where \ttp{property} is a QuickCheck library function that lifts a \ttp{Bool} into
the \ttp{Property} type.  (We wrap \ttp{Int} in a constructor to prevent QuickCheck
from using a default implementation of \ttp{shrink} for \ttp{Int}.)  A typical
counterexample returned by QuickCheck is
%
\begin{code}
> [A (-2),A (-9),A (-26),A (-28),A (-22)]  
\end{code}
%
\noindent
Whereas the smallest counterexample is
%
\begin{code}
> [A 0,A 0,A 0,A 0,A 0]
\end{code}
%
\noindent
The problem is that while QuickCheck is generating larger lists, it is also
generating larger integers.

SmallCheck is another testing framework that is guaranteed to return the a
smallest counterexample, if one exists~\cite{sc}.  SmallCheck does this by
enumerating all possible inputs, ordered from smallest to largest, up to some
user-defined bound.  While SmallCheck is effective for testing many programs and
properties (in accordance with the small model principle~\cite{jackson}),
counterexamples deep in the state space may be infeasible to uncover.  For
example, consider again our data type \ttp{A} defined above, used in the
following property:
%
\begin{code}
data B = B A A A A deriving Show

prop :: B -> Bool
prop (B a b c d) = and $ map f [a, b, c, d]
  where f (A x) = x < 100
\end{code}
%
\noindent
With SmallCheck, a property is checked up to a depth set by the user, measuring
how large values are.  For example, for type \ttp{Int}, the depth \ttp{d} is
defined as the integers in the list \ttp{(-d)..d]}.  For a product type,
  SmallCheck generates the cross product of its fields.  In the example above,
  SmallCheck (or Lazy SmallCheck) will find a counterexample at depth 101,
  requiring approximately $100^4$ test cases, which is infeasible to compute.\footnote{This
  is an upper-bound, since Lazy SmallCheck eliminates some redundant test-cases,
  but computing a counterexample is still infeasible.}

quantified away

and straightforward
implementations of shrink are ineffective.

writing good shrink functions is tedious, difficult, or
both.

In principle, for a given counterexample, \ttp{shrink} can be implemented to
find a local minimum within the neighborhood of the counterexample.

One of the
dirty little secrets of QuickCheck is that even for simple data types and simple
properties, counterexamples can be large, and straightforward implementations of
\ttp{shrink} can miss finding minimal counterexamples, by a lot.

%% For example, consider a trivial arithmetic expression language, containing
%% constants, addition, and division:
%% %
%% \begin{code}
%% data Exp = C Int
%%          | Add Exp Exp
%%          | Div Exp Exp
%% \end{code}
%% %
%% \noindent
%% And consider the following evaluation function that returns \ttp{Nothing} if the
%% expression contains division by zero and

%% \todo{Example}

\todo{then explain why---need to understand property}

\todo{SmallCheck is guaranteed to find minimal counterexamples.}

\todo{insight: shrink based on arbitrary}

\url{http://stackoverflow.com/questions/8788542/how-do-i-get-good-small-shrinks-out-of-quickcheck}

@@

\todo{\cite{syb} paper is motivated by generic shrink}
\todo{shrink paper rehger pointed to ``Automatic isolation of compiler errors'' from 1994}

@@

\section*{Acknowledgements}
\todo{Thank Rehger for comments/pointing to paper.}
